\chapter{Understanding LSTM Locomotion Mode Recognition Behaviour}
\label{chp:lstm-general}

\section{Introduction and Commentary}
% Where possible please ensure that the page numbers tie in with the rest of the thesis document, 
% where this is not possible, you should insert an extra page before the academic paper which shows 
% the citation details and gives the page numbers of the thesis that the academic paper covers.
A large issue identified in literature was the lack of performance classification of novel unseen subjects. To address this research gap further work is required on understanding what prevents \acrshort{lstm} networks from correctly classifying novel subjects.

The \acrshort{lstm} network was selected as HAR is a time-series classification problem with, where input data that are close in space may be dependent while distant sequence of samples in time are assumed as independent. Rae et al.~showed that the \acrshort{lstm} networks adjust well to subject-specific variations\cite{Rai2019}. Thus, the LSTM is renowned for its performance on \acrshort{lmr} tasks\cite{Ordonez2016, Cheng2019, Pienaar2019} 

Following the data collection, work began on developing a general-purpose \acrfull{lstm} network for classifying the locomotive mode of a previously unseen user. This involved work to understand how LSTM networks can recognise different locomotive modes. Once completed, an article entitled ``Understanding LSTM Network Behaviour of IMU-Based Locomotion Mode Recognition for Applications in Prostheses and Wearables'' was submitted to the Journal Sensors. The paper was submitted as part of a call for papers on Machine Learning and Multimodal Sensing for Smart Wearable Assistive Robotics and published on the 10\textsuperscript{th} of February 2021.

The remainder of this chapter will be the presentation of the paper, followed by a short post-commentary. The paper is re-typeset to match the thesis. Page, section and reference numbering has been changed from the published copy.

%\newcommand\x{\value{page}}
 %The paper is presented as published and as such contains both the journal and thesis page numbering. The full paper covers thesis pages \the\numexpr\x+2\relax \ to \the\numexpr\x+24\relax.


\section*{Authorship and permissions}
% \newcommand{\backgoundColor}{0.871,0.918,0.965}
\newcommand{\backgoundColor}{1,1,1}
\newcommand{\boxColor}{0.95,0.95,0.95}
\begin{table}[H]
\caption{Statement of Authorship}
\hyphenpenalty=10000
\begin{tabularx}{\textwidth}{>{\raggedright}p{1.5in}X}
\hline
\rowcolor[rgb]{\backgoundColor}
\multicolumn{2}{l}{\textbf{This declaration concerns the article entitled:}}\\
\hline
\multicolumn{2}{p{0.9\textwidth}}{Understanding LSTM Network Behaviour of IMU-Based Locomotion Mode Recognition for Applications in Prostheses and Wearables} \\

\hline
\rowcolor[rgb]{\backgoundColor}
\multicolumn{2}{l}{\textbf{Publication status (tick one)}}  \\
\rowcolor[rgb]{\backgoundColor}
\multicolumn{2}{l}{\begin{tabular}{rcrcrcrcrc}
    \textbf\ {Draft} &{\cellcolor[rgb]{\boxColor}} \ &
    \textbf{\ Submitted} &{\cellcolor[rgb]{\boxColor}} \ &
    \textbf{\ In review} &{\cellcolor[rgb]{\boxColor}} \ &
    \textbf{\ Accepted} &{\cellcolor[rgb]{\boxColor}} \ &
    \textbf{\ Published} &{\cellcolor[rgb]{\boxColor}}X 
\end{tabular}}  \\
\rowcolor[rgb]{\backgoundColor}\multicolumn{2}{l}{} \\

% Publication details
\hline
{\cellcolor[rgb]{\backgoundColor}}\textbf{Publication details (reference)} & Sensors 2021, 21, 1264.

DOI: \href{https://doi.org/10.3390/s21041264}{10.3390/s21041264} 

Received: 23 December 2020, Accepted: 6 February 2021, Published: 10 February 2021\\


\hline
\rowcolor[rgb]{\backgoundColor}
\multicolumn{2}{l}{\textbf{Publication status (tick one)}} 

\\
\rowcolor[rgb]{\backgoundColor}
\multicolumn{2}{l}{\begin{tabular}{rcrc}
    \textbf{I hold the copyright} &{\cellcolor[rgb]{\boxColor}} \ &
    \textbf{\ Copyright is retained by the publisher, but} &{\cellcolor[rgb]{\boxColor}}X \\
    \textbf{for this material}& & \textbf{I have been given permission to replicate} & \\
    & & \textbf{ the material here} & \\
\end{tabular}}  \\
\rowcolor[rgb]{\backgoundColor}\multicolumn{2}{l}{} \\

\hline
{\cellcolor[rgb]{\backgoundColor}}\textbf{Candidate’s contribution to the paper (provide details, and also indicate as a percentage)} & The candidate predominantly executed the:

\textbf{Formulation of ideas: (95\%)}

The experiment and analysis methodology was conceived by Freddie Sherratt with supervision from Pejman Iravani

\textbf{Design of methodology: (95\%)}

The experiment and analysis methodology was conceived by Freddie Sherratt with supervision from Pejman Iravani

\textbf{Experimental work: (95\%)}

The experimental work was carried out entirely by Freddie Sherratt with supervision from Pejman Iravani

\textbf{Presentation of data in journal format: (95\%)}

The data presentation was carried out entirely by Freddie Sherratt with supervision from Pejman Iravani and Andrew Plummer\\

\hline
{\cellcolor[rgb]{\backgoundColor}}\textbf{Statement from Candidate} &
This paper reports on original research I conducted during the period of my Higher Degree by Research candidature.\\

\hline
{\cellcolor[rgb]{\backgoundColor}}\textbf{Signed} & Freddie Sherratt 
{\begin{tabular}{p{0.6in}ll}
     \ & {\cellcolor[rgb]{\backgoundColor}}\textbf{\ \ Date } & 4th March 2021 
\end{tabular}}\\
\hline
\end{tabularx}
\end{table}
\clearpage

% Paper
% \phantomsection
% \addtocounter{section}{1}
% \addcontentsline{toc}{section}{\arabic{chapter}.\arabic{section}~~~ Article: Understanding LSTM Network Behaviour of IMU-Based Locomotion Mode Recognition  for Applications in Prostheses and Wearables}

% This is the published PDF copy
% \includepdf[pages=-,scale=.83,offset=11mm 8mm,pagecommand={\thispagestyle{fancy}}, frame=false]{content/4-LSTM_Behaviour/sensors-21-01264-v2.pdf}

% This is a re-typeset of the paper
\input{content/4-Paper_Understanding_LSTM}
\clearpage


\section{Post-commentary}
Since being published, seventeen papers have cited the article\footnote{as of the 23\textsuperscript{rd} July 2022}. A number of these citations have used the paper as an example use case for an \acrshort{lstm} network\cite{Uddin2021, Du2021, Velezguerrero2021, Low2022, Bittibssi2022} and data capture methods\cite{Shin2021, Su2021}. The paper was also awarded a Editor's Choice award. The award is given to articles the journal editors believe are of particularly interesting or important in a field.

Maximum performance for validation data was 96.1\% . This classification accuracy is comparable in performance with literature. Uddin et al.~achieved 94\%\cite{Uddin2021} and Murad et al.~achieved 97\%\cite{Murad2017}. However the performance of the test set, containing novel unseen individuals, was significantly lower at 76\%. 

This results supports the conclusion drawn by Dehghani et al. that a substantial drop in performance for unseen subjects is to be expected\cite{Dehghani2019}. This work extends on that conclusion by proposing that the high variability between individuals during early stance is a large source of this error. As when the target user behaviour differs from the training dataset the subject will perform poorly\cite{Qiu2022}. 

Therefore for a general model to perform well the training set would need to include individuals with a similar gait pattern to all target subjects. This is highly impractical especially for amputee who have significantly more variability in their gait\cite{Lechler2018, Kovac2009}. It is also harder to get data for amputees due to reduced mobility therefore harder to add them into the general training pool\cite{Gardiner2016}.

% What was the big takeaway from this paper - demonstrates the need for personalisation
The paper's outcomes expressed the need to adapt the trained model to each target individual through a method of personalisation in order to perform adequately. It also noted the need for additional testing to demonstrate the suitability of these techniques for amputees. These needs will be explored further in the subsequent chapters.

